An Empirical Analysis of the Internet Engineering Task Force with Computational Methods

### 2024-10-06 paper.tex
Balance columns
Related work

\documentclass[twocolumn,10pt]{article}
\pdfoutput=1 % This must be in the first 5 lines to tell arXiv to use pdfLaTeX
\usepackage[l2tabu,orthodox]{nag}
\usepackage[utf8x]{inputenc}
\usepackage[british]{babel}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage[all]{onlyamsmath}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{upquote}
\usepackage{graphicx}
\usepackage{url}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{color}
\usepackage[cm]{fullpage}
\usepackage{multirow}
\usepackage{authblk}
\usepackage{hyperref}

@@

@@

%==================================================================================================
\bibliographystyle{abbrv}
\bibliography{paper}

>>>>

\documentclass[twocolumn,10pt]{article}
\pdfoutput=1 % This must be in the first 5 lines to tell arXiv to use pdfLaTeX
\usepackage[l2tabu,orthodox]{nag}
\usepackage[utf8x]{inputenc}
\usepackage[british]{babel}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage[all]{onlyamsmath}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{upquote}
\usepackage{graphicx}
\usepackage{url}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{color}
\usepackage[cm]{fullpage}
\usepackage{multirow}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{balance}

@@

%..................................................................................................
% IMC 2021 paper:

To the best of our knowledge, were the first to perform a comprehensive
statistical analysis of IETF activities considering trends in document
publication (\S\ref{sec:trends-documents}), author demographics
(\S\ref{sec:trends-demographics}), and communication patterns.
\S\ref{sec:org-dyn:participants}. That said, there have been several
prior efforts that have focused on specific aspects of the IETF:

\begin{itemize}
  \item Jari Arkko, a former Chair of the IETF, maintains a website
    (\url{https://www.arkko.com/tools/stats.html}) that provides various
    statistics about the IETF, including about its documents, authors and
    their affiliations. The Datatracker also provides some statistics on
    documents and authors (\url{https://datatracker.ietf.org/stats/}).
    Attendance statistics are reported in the IETF plenaries (available
    in the proceedings at \url{https://www.ietf.org/meeting/past/}).

  \item BigBang~\cite{benthall2015testing} is a Python toolkit for
    analysing online collaborative communities through mailing list data.
    Niedermayer et al.~\cite{niedermayer2017information} discuss the
    challenges of working with large mailing list datasets, including
    conducting  entity resolution where contributors have changeable
    pseudonyms. Our work addresses a number of these challenges by
    augmenting the mailing list data with metadata from the IETF
    Datatracker.

  \item Huitema~\cite{RFC8963} carried out an evaluation of a small subset
    of RFCs, to understand the various sources of delay in their
    publication. The author observed that the main source of delay was the
    working group process, highlighting the need to better understand the
    dynamics of that process. In addition, the author found that citation
    counts did not correlate well with the adoption or deployment of
    specific RFCs.
\end{itemize}

%..................................................................................................
% ICWSM 2022 paper:

Our work to characterise participant influence and impact (\S\ref{sec:org-dyn:influence})
builds on a rich literature regarding influence in online social networks
and other communities:
\begin{itemize}
  \item 
    Studies of Twitter~\cite{cha2010measuring,weitzel2012measuring,anger2011measuring},
    Instagram~\cite{zarei2020characterising} and decentralised social networks
    \cite{harissigmetrics, hassanpleroma}. For example, \cite{ye2010measuring}
    measured the impact of follower influence on message propagation. Others
    have focused on devising metrics to capture influence: similarly
    to~\cite{kourtellis2013identifying,weitzel2012measuring,sole2014centrality},
    we use betweenness centrality as a metric of influence.

  \item Online collaborative communities such as Wikimedia and Open Stack
    have also been studied for understanding communities, collaboration,
    strategy making, and organisational structures, through mailing lists
    and platform interactions \cite{dobusch2018open, bosu2014social,
    zhang2020companies}. Bosou and Carver \cite{bosu2014impact} highlight
    that contributors with much higher reputation are successful in seeking
    reviews from the community in a much shorter time span and are more
    likely to get their suggestions accepted. This is well aligned with
    most of our findings related to the impact of influential people on the
    Internet-Draft adoption process. Zhang et al.\ observe
    \cite{zhang2020companies} that in the open source software (OSS)
    ecosystem, several participating organisations (firms or companies) may
    engage in intentional or passive collaborations, or they may contribute
    in an isolated way. They find that an organisation's influence in the
    collaboration network is positively correlated with its scale of
    contribution within the ecosystem. In some ways, this is similar to our
    observation related to prominent affiliated organisations in the
    co-authorship network and their ability to collaborate and produce
    drafts.

  % This bullet adapted from our ICWSM 2024 paper
  \item Building social interaction networks out of email communication
    data is a well trodden field~\cite{belanger1999communication,
    mcpherson2001birds}, and the introduction of time evolution into the
    analysis~\cite{panzarasa2009patterns,viswanath2009evolution} is recent
    but has quickly gained interest. However, the analysis of email based
    communication is less prevalent with the notable exception of the
    well-known, and oft analysed, Enron emails~\cite{klimt2004enron}. The
    communication dataset we examine is interesting for several reasons: it
    is a long duration email dataset, it is annotated with meta data
    representing participant roles, and it represents discussion in a
    decentralised consensus driven organisation~\cite{RFC7282,RFC3935}
    rather than a business with a traditional management hierarchy.
    While the Enron data set is often used for classification of emails
    into categories \cite{madjarov2012extensive}, there are also several
    studies of organisation interaction patterns on it.  Namely,
    \cite{tang2008community,kossinets2008structure} are concerned with
    developing methodologies for studying how communication evolves over
    time.  In contrast, \cite{diesner2005communication} focus on providing
    insights more than on developing a methodology, revealing that
    organisational structure is reflected in  email communication patterns.
    Other studies focusing on mailing list data include OSS mailing lists
    \cite{bird2006mining,rigby2007can}.

\end{itemize}

%..................................................................................................
% ICWSM 2024 paper

Analysing the temporal patterns in communication, as we do for the IETF
organisational hierarchy in \S\ref{sec:org-dyn:hierarchy}, is an emerging
field of research~\cite{masuda2016guide}:

\begin{itemize}
  \item Some have analysed the time evolution of important nodes in a
    network~\cite{fire2020rise}, whether their neighbours are similarly
    important~\cite{pedreschi2022temporal} and some others the correlations
    of the trajectory of snapshots of networks over
    time~\cite{lacasa2022correlations}. In this paper we bring new analysis
    techniques, that expand on the idea of tracking the evolution of node
    importance, to an email list corpus.

  \item A form of temporal network analysis that is key to understanding
    communication in social networks is computing temporal
    motifs~\cite{paranjape2017motifs,kovanen2011temporal} between nodes.
    This technique involves look at the sequence in which edges attach to
    nodes over time which lets us see how conversation happens with fine
    resolution.  Applications include understanding the diffusion of
    information through a social network using undirected
    motifs~\cite{sarkar2019using}, and analysing the collaboration and
    scientific mobility of co-authorship
    networks~\cite{boekhout2021investigating}. To the best of our
    knowledge, temporal motifs have not yet been used for analysis of
    communication in social networks which means knowledge of high
    time-resolution direct communication between individuals is ripe for
    the picking.

  \item The effects of hierarchy on the ``voice" of individuals within an
    organisation is a key aspect~\cite{pfrombeck2023hierarchy} of
    Organisational Behaviour (OB) research. A steep~\cite{hussain2019voice}
    hierarchy can have a negative effect on voice by reducing the variety
    of perspectives~\cite{anderson2010functions} in higher levels, or
    positive effect where decisions can be made quickly. However, a
    diffused hierarchy may also have negative
    effects~\cite{hussain2019voice} on voice by contributing to lower level
    ``bystanders" who let the higher levels do the talking, or a positive
    effect with a variety of approaches to management. Similarly, a large
    ``power distance" can cause problems to lower level voice if the higher
    levels are ``authoritarian"~\cite{duan2018authoritarian}, however if
    they are ``benevolent"~\cite{li2021does,guo2020inclusive} then power
    distance is less of a problem. We aim to test these ideas in regard to
    the voluntary and collaborative nature of the IETF organisation,
    whereas the literature mostly focuses on traditional organisations.

\end{itemize}

%..................................................................................................

% The following is adapted from our ACL 2023 Tracing Linguistic Markers of
% Influence paper:

Considering use of language (\S\ref{sec:language}), most studies of
influence either focus on community structure rather than language, or use
language indirectly.  Urena et al.\ \cite{urena2019review} give a survey of
the former approach. In an example of the latter, Prabhakaran
\cite{prabhakaran2014staying} compares users with different influence in
terms of their linguistic similarity or co-adaptation, the increasing
similarity of interlocutors to each other in how they use language (see
also \cite{danescu2012echoes,ver2013information,noble2015centre,
kawabata2016convex,buske2019linguistic}).  
Some studies focus on modelling influence from text of Enron emails by
identifying keywords/phrases that indicate influence \cite{bramsen2011extracting,
gilbert2012phrases}.  Rosenthal \cite{rosenthal2014detecting} and Tchokni
et al.\ \cite{tchokni2014emoticons} extend this approach to other domains,
including Twitter, Wikipedia talk pages, and debates, and include a wider
range of linguistic markers.

%..................................................................................................
The work in \S\ref{sec:success-factors} builds, in part, on Nikkhah et
al.~\cite{nikkhah2017statistical} by statistically exploring RFC adoption.
We build on this work with improved modelling, and by incorporating
additional features that characterise not only the RFCs themselves but also
the standardisation process that produced them.

@@

%==================================================================================================
\balance
\bibliographystyle{abbrv}
\bibliography{paper}
